Analysis for Biracial Grade Retention Project

### 2022-09-03 analysis/analysis.Rmd
Use survey weighted values in all tables

@@

Lets looks at sample characteristics by race.

```{r desc-stats}
tbl_summary <- acs %>%
  group_by(race) %>%
  summarize(n=n(), 
            held_back=round(mean(below_exp_grade), 3)*100,
            mean_age=round(mean(age), 1),
            f_born=round(mean(foreign_born), 3)*100,
            spk_eng_well=round(mean(spk_eng=="Yes"), 3)*100,
            spk_eng_par=round(mean(spk_eng_father=="Yes" & 
                                     spk_eng_mother=="Yes"), 3)*100,
            par_fb=round(mean(foreign_born_father | 
                                       foreign_born_mother), 3)*100,
            fam_inc=round(mean(family_income), 0),
            own_home=round(mean(own_home), 3)*100,
            par_hs=round(mean(degree_father!="LHS" | 
                                degree_mother!="LHS"), 3)*100,
            par_ba=round(mean(degree_father=="BA" | 
                                degree_father=="G" | 
                                degree_mother=="BA" | 
                                degree_mother=="G"), 3)*100,
            par_mar=round(mean(parents_married), 3)*100) %>%
  arrange(desc(n))

@@

tbl_summary %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Descriptive statistics by race. Shading indicates biracial group. Groups ordered by sample size.",
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race","n","% below expected grade", "mean age",
                  "% foreign-born","% speak English well",
                  "% Both parents speak english well","% either parent foreign-born",
                  "mean family income","% own home","% either parent, HS diploma",
                  "% Either parent, BA degree","% parents married")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

@@

All models are estimated as logit models. Warning, the code chunks below take a bit of time to run!

```{r create-design}
# strata make the models take forever, but do not seem to change anything
# so I will leave them out
acs_design <- svydesign(ids=~cluster, weights=~perwt, data=acs)
```

@@

Its easy to get a simple adjusted prediction at specific values using the `predict` command, but more complicated for the average adjusted prediction. In order to calculate these average adjusted predictions, I use the `marg` command in the `modmargs` library. This function has a nice `type="levels"` argument that directly gives you the average adjusted prediction. Standard errors for these average adjusted predictions are calculated by the [delta method](https://cran.r-project.org/web/packages/modmarg/vignettes/delta-method.html). I also adjust all SEs for clustering of respondents within households.

@@

```{r desc-stats-ses}
tbl_summary <- acs %>%
  group_by(race) %>%
  summarize(fam_inc=round(mean(family_income), 0),
            own_home=round(mean(own_home), 3)*100,
            par_mar=round(mean(parents_married), 3)*100,
            f_lhs=round(mean(degree_father=="LHS"), 3)*100,
            f_hs=round(mean(degree_father=="HS"), 3)*100,
            f_aa=round(mean(degree_father=="AA"), 3)*100,
            f_ba=round(mean(degree_father=="BA" | degree_father=="G"), 3)*100,
            m_lhs=round(mean(degree_mother=="LHS"), 3)*100,
            m_hs=round(mean(degree_mother=="HS"), 3)*100,
            m_aa=round(mean(degree_mother=="AA"), 3)*100,
            m_ba=round(mean(degree_mother=="BA" | degree_mother=="G"), 3)*100)

@@

```{r desc-stats-cult}
tbl_summary <- acs %>%
  group_by(race) %>%
  summarize(f_born=round(mean(foreign_born), 3)*100,
            father_fb=round(mean(foreign_born_father), 3)*100,
            mother_fb=round(mean(foreign_born_mother), 3)*100,
            spk_eng_no=round(mean(spk_eng=="No"), 3)*100,
            spk_eng_some=round(mean(spk_eng=="Somewhat"), 3)*100,
            spk_eng_yes=round(mean(spk_eng=="Yes"), 3)*100,
            f_spk_eng_no=round(mean(spk_eng_father=="No"), 3)*100,
            f_spk_eng_some=round(mean(spk_eng_father=="Somewhat"), 3)*100,
            f_spk_eng_yes=round(mean(spk_eng_father=="Yes"), 3)*100,
            m_spk_eng_no=round(mean(spk_eng_mother=="No"), 3)*100,
            m_spk_eng_some=round(mean(spk_eng_mother=="Somewhat"), 3)*100,
            m_spk_eng_yes=round(mean(spk_eng_mother=="Yes"), 3)*100)

@@

tbl_resources <- acs %>%
  group_by(race) %>%
  summarize(fam_inc=round(mean(family_income), 0),
            own_home=round(mean(own_home), 3)*100,
            m_ba=round(mean(degree_mother=="BA" | degree_mother=="G"), 3)*100,
            f_ba=round(mean(degree_father=="BA" | degree_father=="G"), 3)*100,
            f_born=round(mean(foreign_born), 3)*100,
            par_fb=round(mean(foreign_born_father | 
                                       foreign_born_mother), 3)*100,
            spk_eng_well=round(mean(spk_eng=="Yes"), 3)*100,
            spk_eng_par=round(mean(spk_eng_father=="Yes" & 
                                     spk_eng_mother=="Yes"), 3)*100) %>%
  mutate(race=as.character(race)) %>%
  left_join(temp) %>%
  arrange(held_back_cf)

@@

tbl_resources %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Resource differences by race. Shading indicates biracial group.",
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race","mean family income","% own home",
                  "% mother, BA degree","% father, BA degree",
                  "% foreign-born","% either parent foreign-born",
                  "% speak English well", "% Both parents speak english well",
                  "% held back, counterfactual")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")

>>>>

```{r create-design}
# strata make the models take forever, but do not seem to change anything
# so I will leave them out
acs_design <- acs %>%
  as_survey_design(ids = cluster, weights = perwt)
```

@@

Lets looks at sample size and grade retention by race.

```{r desc-stats}
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(n=n(), 
            held_back=round(survey_mean(below_exp_grade, vartype=NULL), 3)*100,
            mean_age=round(survey_mean(age, vartype=NULL), 1)) %>%
  arrange(desc(n))

@@

tbl_summary %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Descriptive statistics by race. Shading indicates biracial group. Groups ordered by sample size. Survey weights applied.",
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race","n","% below expected grade", "mean age")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

@@

All models are estimated as logit models. I use the `survey` library to adjust for variable sample weights and clustering of respondents in the same household. Warning, the code chunks below take a bit of time to run!

@@

Its easy to get a simple adjusted prediction at specific values using the `predict` command, but more complicated for the average adjusted prediction. In order to calculate these average adjusted predictions, I use the `marg` command in the `modmargs` library. This function has a nice `type="levels"` argument that directly gives you the average adjusted prediction. Standard errors for these average adjusted predictions are calculated by the [delta method](https://cran.r-project.org/web/packages/modmarg/vignettes/delta-method.html). I also adjust all SEs for clustering of respondents within households and variance in person weights.

@@

```{r desc-stats-ses}
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=survey_mean(family_income, vartype = NULL))

tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=round(survey_mean(family_income, vartype = NULL), 0),
            own_home=round(survey_mean(own_home, vartype = NULL), 3)*100,
            par_mar=round(survey_mean(parents_married, vartype = NULL), 3)*100,
            f_lhs=round(survey_mean(degree_father=="LHS", vartype = NULL), 
                        3)*100,
            f_hs=round(survey_mean(degree_father=="HS", vartype = NULL), 3)*100,
            f_aa=round(survey_mean(degree_father=="AA", vartype = NULL), 3)*100,
            f_ba=round(survey_mean(degree_father=="BA" | degree_father=="G", 
                                   vartype = NULL), 3)*100,
            m_lhs=round(survey_mean(degree_mother=="LHS", vartype = NULL), 3)*100,
            m_hs=round(survey_mean(degree_mother=="HS", vartype = NULL), 3)*100,
            m_aa=round(survey_mean(degree_mother=="AA", vartype = NULL), 3)*100,
            m_ba=round(survey_mean(degree_mother=="BA" | degree_mother=="G", 
                                   vartype = NULL), 3)*100)

@@

```{r desc-stats-cult}
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(f_born=round(survey_mean(foreign_born, vartype = NULL), 3)*100,
            father_fb=round(survey_mean(foreign_born_father, 
                                        vartype = NULL), 3)*100,
            mother_fb=round(survey_mean(foreign_born_mother, 
                                        vartype = NULL), 3)*100,
            spk_eng_no=round(survey_mean(spk_eng=="No", vartype = NULL), 3)*100,
            spk_eng_some=round(survey_mean(spk_eng=="Somewhat", 
                                           vartype = NULL), 3)*100,
            spk_eng_yes=round(survey_mean(spk_eng=="Yes", vartype = NULL), 3)*100,
            f_spk_eng_no=round(survey_mean(spk_eng_father=="No", 
                                           vartype = NULL), 3)*100,
            f_spk_eng_some=round(survey_mean(spk_eng_father=="Somewhat", 
                                             vartype = NULL), 3)*100,
            f_spk_eng_yes=round(survey_mean(spk_eng_father=="Yes", 
                                            vartype = NULL), 3)*100,
            m_spk_eng_no=round(survey_mean(spk_eng_mother=="No", 
                                           vartype = NULL), 3)*100,
            m_spk_eng_some=round(survey_mean(spk_eng_mother=="Somewhat", 
                                             vartype = NULL), 3)*100,
            m_spk_eng_yes=round(survey_mean(spk_eng_mother=="Yes", 
                                            vartype = NULL), 3)*100)

@@

tbl_resources <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=round(survey_mean(family_income, vartype=NULL), 0),
            own_home=round(survey_mean(own_home, vartype=NULL), 3)*100,
            m_ba=round(survey_mean(degree_mother=="BA" | degree_mother=="G", 
                                   vartype=NULL), 3)*100,
            f_ba=round(survey_mean(degree_father=="BA" | degree_father=="G", 
                                   vartype=NULL), 3)*100,
            f_born=round(survey_mean(foreign_born, vartype=NULL), 3)*100,
            par_fb=round(survey_mean(foreign_born_father | foreign_born_mother, 
                                     vartype=NULL), 3)*100,
            spk_eng_well=round(survey_mean(spk_eng=="Yes",
                                           vartype=NULL), 3)*100,
            spk_eng_par=round(survey_mean(spk_eng_father=="Yes" & 
                                     spk_eng_mother=="Yes", vartype=NULL), 
                              3)*100) %>%
  mutate(race=as.character(race)) %>%
  left_join(temp) %>%
  arrange(held_back_cf)

@@

tbl_resources %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(caption="Resource differences by race. Shading indicates biracial group. All results are survey weighted.",
      align = c("l",rep("r", ncol(tbl_resources)-1)),
      col.names=c("race","mean family income","% own home",
                  "% mother, BA degree","% father, BA degree",
                  "% foreign-born","% either parent foreign-born",
                  "% speak English well", "% Both parents speak english well",
                  "% held back, counterfactual")) %>%
  kable_paper() %>%
  row_spec(index_biracial, background = "#D3D3D3")
